### 2014-07-25 Paper_v1.tex
Updated text and figures

>>>>

\subsection*{Inhibition of Raf kinase activity modulated MAPK activation and differentiation markers.}
Next, we considered the effect of the Raf kinase inhibitor GW5074 on functional markers of ATRA-induced growth arrest and differentiation.
For more details on GW5074 including dose response see Supplemental Materials including Fig. \ref{F:MAPKGWWB}A.
ATRA treatment alone statistically significantly increased the G1/G0 percentage over the untreated control, 
while GW5074 alone had a negligible effect on the cell cycle distribution (Fig. \ref{F:GWdiff}A).
Secondly, ATRA treatment alone increased both Raf phosphorylation at S621/S259 and Raf expression 24 hr post-treatment,
while ATRA-induced phosphorylation of ERK was blocked by GW5074, consistent with the loss of Raf kinase activity (Fig. \ref{F:MAPKGWWB}B).
Lastly, both the phosphorylation and expression of Raf was inhibited by GW5074.
This suggested a positive feedback between Raf kinase activity, Raf phosphorylation and ERK action on Raf transcription factors.
Surprisingly, the combination of GW5074 and ATRA statistically significantly increased the G1/G0 population (82 $\pm$ 1\%)
compared with ATRA alone (61 $\pm$ 0.5\%).
Increased G1/G0 arrest following the combined treatment with GW5074 and ATRA was unexpected, 
as the combination of ATRA and the MEK inhibitor (PD98059) has been shown previously to decrease ATRA-induced growth arrest \cite{Yen1998}.
However, growth arrest is not the sole indication of functional differentiation.
Expression of the cell surface marker CD11b has also been shown to coincide with HL-60 cells myeloid differentiation \cite{Hickstein1989}.  
We measured CD11b expression, for the various treatment groups, using immuno-fluorescence flow cytometry 48 hr post-treatment.
As with G1/G0 arrest, ATRA alone increased CD11b expression over the untreated control, 
while GW5074 further enhanced ATRA-induced CD11b expression (Fig. \ref{F:GWdiff}B).
GW5074 alone had no statistically significant effect on CD11b expression, 
compared with the untreated control.

The inducible reactive oxygen species (ROS) response was used as a functional marker of differentiated neutrophils \cite{Congleton2011}.
We measured the ROS response induced by the phorbol ester 12-O-tetradecanoylphorbol-13-acetate (TPA) using flow cytometry.
Untreated cells showed no discernible TPA response, 
with only 7.0 $\pm$ 3.0\% ROS induction (Fig. \ref{F:GWdiff}C).
Cells treated with ATRA had a significantly increased TPA response, 53 $\pm$ 7\% ROS induction 48 hr post-treatment.
Treatment with both ATRA and GW5074 statistically significantly reduced ROS induction (22 $\pm$ 0.6\%) compared to ATRA alone. 
Interestingly, Western blot analysis did not detect a GW5074 effect on ATRA-induced expression of p47phox, 
a required upstream component of the ROS response  (Fig. \ref{F:GWdiff}C, bottom).
Thus, the inhibitory effect of GW5074 on inducible ROS might occur downstream of p47phox expression.
However, the ROS producing complex is MAPK dependent, 
therefore it is also possible that GW5074 inhibited ROS production by interfering 
with MAPK activation (in which case the p47phox marker might not accurately reflect phenotypic conversion and differentiation).
